Models, derivatives, response diversity

### 2024-07-08 models_derivatives_response_diversity.Rmd
re-testing the addition of lepomis gibbosus

# # Model 4: before simplification
# species_list <- df_abundance_re |>
#   # no simplification
#   filter(Species %in% c("Alburnus_arborella", "Barbatula_sp_Lineage_I",
#                         "Cyprinus_carpio", "Phoxinus_csikii", "Salmo_trutta")) |>
#   distinct(Species) |> 
#   pull(Species)
# 
# species_list <- sort(species_list)
# 
# derivatives <- list()
# gam_output <- list()
# model_prediction <- list()
# 
# 
# for (i in species_list) {
#   
#   data <- df_abundance_re |> 
#     filter(Species == i)
#   
#   gam_output[[i]] <- gam(data = data, Abundance ~ s(mean_last_7days, k = 3) + s(fLake, bs = 're')
#                          +  s(fProtocol, bs = 're'), family = ziP())
#   
#   lake_data <- species_lake |>
#     filter(Species == i)
#   
#   lake_list <- distinct(lake_data, Lake) |>
#     pull()
#   
#   unique_lakes <- distinct(data, Lake) |> 
#     pull()
#   
#   for (j in lake_list){
#     
#     data_lake <- species_lake |>
#       filter(Species == i) |>
#       filter(Lake == j)
#    
#     random_lake <- sample(unique_lakes, 1)
# 
#     newdata <- tibble(mean_last_7days = seq(
#       from = min(data_lake$temp, na.rm = TRUE),
#       to = max(data_lake$temp, na.rm = TRUE), length = 200),
#       fLake = factor(random_lake), fProtocol = factor("VERT"))
#     
#     derivatives <- derivatives(gam_output[[i]], data = newdata) |>
#       mutate(fLake = factor(j)) |>
#       mutate(species = factor(i)) |>
#       rename(temp = mean_last_7days)
#     
#     saveRDS(derivatives, paste0("total_models/derivatives_no_simplification/derivatives_", i, "_",  j, ".rds"))
# 
#   }
# }

# save total derivatives as RDS (no exclusions)
# 
# df_deriv_no_excl <- list.files(path = "total_models/derivatives_no_simplification", pattern = ".rds", full.names = TRUE) |>
#   map_dfr(readRDS)
# 
# saveRDS(df_deriv_no_excl, "total_models/df_deriv_all_no_excl.rds")

>>>>

# Model 4: before simplification
species_list <- df_abundance_re |>
# #   # no simplification
filter(Species %in% c("Alburnus_arborella", "Barbatula_sp_Lineage_I",
                     "Cyprinus_carpio", "Phoxinus_csikii", "Salmo_trutta")) |>
distinct(Species) |> 
pull(Species)

for (i in species_list) {
 data <- df_abundance_re |> 
     filter(Species == i)

  for (j in lake_list){
    data_lake <- species_lake |>
      filter(Species == i) |>
     filter(Lake == j)

     saveRDS(derivatives, paste0("total_models/derivatives_no_simplification/derivatives_", i, "_",  j, ".rds"))

# save total derivatives as RDS (no exclusions)
#
df_deriv_no_excl <- list.files(path = "total_models/derivatives_no_simplification", pattern = ".rds", full.names = TRUE) |>
  map_dfr(readRDS)
#
saveRDS(df_deriv_no_excl, "total_models/df_deriv_all_no_excl.rds")
